Paper

### 2019-11-21 Paper_resubmit.Rmd
...

p11Loss %>% filter(value < -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% full_join(
  p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
  kable(caption='Chr17 p11 loss') %>% kable_styling(full_width = F)

p11Loss %>% group_by(Status, Hospital.Research.ID) %>% tally

@@

```{r}
patient.info %>% filter(Samplename %in% p11Loss$variable) %>% group_by(forcats::fct_explicit_na(p53.Status)) %>% tally %>% 
  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 status for samples with p11 loss') %>% kable_styling(full_width = F)

@@

cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
```

@@

## Chr 17 Predictive TODO

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
  filter(!grepl('^X|Y',`Genome Location`)) %>% 
  separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
  dplyr::filter(start != '' | end != '') %>%
  dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)

genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) | `Mutation Types` != 'T') %>% 
  dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
  dplyr::mutate(value = min(globalCNdf$value), Patient = '', chr = factor(chr,levels=levels(chr.lengths$chr),ordered=T), start = as.numeric(start), end=as.numeric(end)) %>% arrange(chr, start)

@@

g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
  labs(title="Chromosome 17") + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
  theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))

@@

for (chr in unique(features$chr)) {  
  print(chr)
  p = plot.cn.chr(globalCNdf,  chrom=chr, info=chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) +
    scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
    theme(strip.text.x = element_blank(), text=element_text(size=14)) + labs(title=paste('Chromosome',chr))
  f = paste('chr',chr,'.png',sep='')
  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=8, height=4, units='in', dpi=300)
}

@@

val.info = readxl::read_xlsx(patient.file, 'Final Validation Samples') %>% 
  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`) %>%
  dplyr::select(-matches('SLX|Index'), -`Path Notes`) %>%
  mutate(Status = factor(Status, levels=levels(pg.samp$Status))) %>% ungroup %>% group_by(`Hospital Research ID`) %>% 
  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Last Endoscopy`, `Endoscopy`, units='weeks')))

@@

@@

val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% dplyr::select(Patient, everything())

val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
  mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))

@@

@@

pt = 34
p34 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34, width=6.35, height=4.5,dpi=72)

pt = 3
p3 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=72)

pt = 68
p68 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=5,dpi=72)

@@

rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  r
}))

rules$t1 = factor(rules$t1, levels=levels(pg.samp$brks), ordered=T)
rules$t2 = factor(rules$t2, levels=levels(pg.samp$brks), ordered=T)

penult = filter(rules, t2 != 'Endpoint' & (t1 == rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))

@@

penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(penult %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Penult Endoscopy: % patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)

endpoint = subset(rules,  t2 != 'Endpoint') %>% 
  dplyr::group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

endpoint %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(endpoint %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (P) per endoscopy') %>% kable_styling(full_width=F)

>>>>

p11Loss %>% filter(value <= -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% 
  full_join( p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
  kable(caption='Chr17 p11 loss') %>% kable_styling(full_width = F)

p11Loss %>% filter(value <= -1) %>% dplyr::select(Patient, Status) %>% distinct %>%
  group_by(Status) %>%  tally(name='p11 Loss') %>%
  kable(caption='Patients showing p53 loss & p53 IHC') %>% kable_styling(full_width = F)

@@

```{r}
patient.info %>% filter(Samplename %in% p11Loss$variable) %>% group_by(forcats::fct_explicit_na(p53.Status)) %>% tally %>% 
  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 IHC status for samples with p11 loss') %>% kable_styling(full_width = F)

@@

ct = cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
```

17p11 loss and p53 IHC correlation  R=`r round(ct$estimate,2)`.

@@

## Chr 17 Predictive

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=T}
ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
  filter(!grepl('^X|Y',`Genome Location`)) %>% 
  separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
  dplyr::filter(start != '' | end != '') %>%
  dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)

# leave out the translocations
genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) & `Mutation Types` != 'T') %>% 
  dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
  dplyr::mutate(value = min(globalCNdf$value), Patient = '', chr = factor(chr,levels=levels(chr.lengths$chr),ordered=T), start = as.numeric(start), end=as.numeric(end)) %>% arrange(chr, start)

@@

g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
  labs(title="Chromosome 17") + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
  theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))
g17

@@

for (chr in unique(features$chr)) {  
  #print(chr)
  p = plot.cn.chr(globalCNdf,  chrom=chr, info=chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) +
    scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
    theme(strip.text.x = element_blank(), text=element_text(size=14)) + labs(title=paste('Chromosome',chr))
  f = paste('chr',chr,'.png',sep='')
  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=8, height=4, units='in', dpi=300)
}

@@

val.info = readxl::read_xlsx(patient.file, 'Final Validation Samples') %>% 
  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`) %>%
  dplyr::select(-matches('SLX|Index'), -`Path Notes`) %>%
  mutate(Pathology = recode(Pathology, 'GM'='NDBE'), Status = factor(Status, levels=levels(pg.samp$Status))) %>% 
  ungroup %>% group_by(`Hospital Research ID`) %>% 
  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Last Endoscopy`, `Endoscopy`, units='weeks')))

@@

oac.preds %>% group_by(`Hospital Research ID`) %>% 
  dplyr::summarise( Probability = max(Probability), `Relative Risk` = max(`Relative Risk`), Risk = max(Risk)) %>%
  group_by(Risk) %>% tally %>% kable(caption='OAC samples used as a positive control') %>% kable_styling(full_width = F)

@@

val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% 
  dplyr::select(Patient, everything()) %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD'), ordered=T)) %>% 
  mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))

@@

# Risk Categories KM

```{r, fig.width=6, fig.height=8}
require(survival)
require(survminer)
require(GGally)

ps = pg.samp %>% dplyr::select(Patient, Status,months.before.final, Risk) %>%
  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)

sfit = survfit( Surv(months.before.final, Diagnosis, type='right') ~ Risk, data=ps, conf.int=0.95, conf.type='log', na.action=na.omit)
names(sfit$strata) = sub('.*=','', names(sfit$strata))

kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
                      conf.int=T, censor=T, legend='none', title='Progression risk over time', xlab='Time (months prior to progression)',
                      ylab = 'Progression risk',
                      ggtheme=theme_minimal(), palette=as.vector(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))

kmp

#ggsave('plots/risk_km.png', plot=print(kmp), height=9, width=8, units = 'in', dpi = 300)

```{r}
vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% mutate(Patient = as.factor(Patient)) %>% 
  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)

merged = bind_rows(ps,vp)

sfit = survfit( Surv(months.before.final, Diagnosis, type='right') ~ Risk, data=merged, conf.int=0.95, conf.type='log', na.action=na.omit)
names(sfit$strata) = sub('.*=','', names(sfit$strata))

kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
                      conf.int=T, censor=T, legend='none', title='Progression Risk', 
                      subtitle='Training & Validation predictions', xlab='Time (months prior to progression)',
                      ylab = 'Progression risk',
                      ggtheme=theme_minimal(), palette=as.vector(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))

kmp

ggsave('plots/risk_km.png', plot=print(kmp), height=9, width=8, units = 'in', dpi = 300)

@@

pt = 34
p34 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34, width=6.35, height=4.5,dpi=72)

pt = 3
p3 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=72)

pt = 68
p68 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=5,dpi=72)

@@

pt.counts = maxP %>% group_by(Patient, Status) %>% dplyr::filter( length(which(grepl('Endpoint', brks))) > 0  & length(brks) > 1) %>% 
  ungroup %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally

rules = do.call(bind_rows, purrr::map(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  return(as_tibble(r))
})) %>% dplyr::mutate(t1 = factor(t1, levels=levels(pg.samp$brks), ordered=T), t2 = factor(t2, levels=levels(pg.samp$brks), ordered=T)) %>% arrange(Patient, t2)

rules %>% group_by(Patient, Status) %>% dplyr::filter(t2 == 'Endpoint' & length(t1) > 2) %>% 
  dplyr::select(Patient, Status) %>% distinct %>%  group_by(Status) %>% tally

ids = rules %>% group_by(Patient, Status) %>% dplyr::filter(t2 == 'Endpoint' & length(t1) > 2) %>% dplyr::select(Patient) %>% pull

penult = rules %>% dplyr::filter(Patient %in% ids & t2 != 'Endpoint') %>% arrange(Patient, desc(t1)) %>% group_by(Patient, Status) %>% 
  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)

#penult = filter(rules, t2 != 'Endpoint' & (t1 %in% rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))

penult %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally

@@

penult %>% filter(Status == 'NP' & rule == 1) %>% kable(caption='NP patients for RFA') %>% kable_styling(full_width = F)

penult %>% group_by(Status) %>% tally(name='total')

penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  kable(caption='Penult Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)

penult = rules %>% dplyr::filter(Patient %in% ids & t2 != 'Endpoint') %>% arrange(Patient, desc(t1)) %>% group_by(Patient, Status) %>% 
  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
